---
title: "Workflow : An Empirical Saddlepoint Approximation Based Method for Smoothing Survival Functions Under Right Censoring"
author:
  name: Pratheepa Jeganathan
  affiliation: Department of Statistics, Stanford University
  email: jpratheepa31@gmail.com
date: "`r Sys.Date()`"
output: 
    rmarkdown::html_vignette:
        toc_float:  true
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Install [R](https://www.r-project.org/) and [RStudio](https://www.rstudio.com/). Open this `Rmd` file in RStudio. Then run the following code to install all required packages.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
pkgs <- c("actuar", "survival", "ggplot2", "gridExtra", "asaur")

#   installed packages that were not installed already
source("http://bioconductor.org/biocLite.R")
biocLite(setdiff(pkgs,installed.packages()), suppressUpdates = TRUE)

devtools::install_github("PratheepaJ/ESPA")
```


```{r}
library(actuar)#     heavy tailed distributions
library(survival)
library(ggplot2)
library(gridExtra)
library(asaur)# dataset "gastricXelox"
library(ESPA)
```


## AML data 

-  Choose the first 11 observations that fall into 'mainitained' group.

```{r}
x <- citation("asaur")
x <- citation("survival")
toBibtex(x)
```

- Data

```{r}
d1 <- aml[1:11,]
time1 <- d1[,1]
status1 <- d1[,2]
```

- KM

```{r}
survob <- Surv(time=time1,event=status1)
fit1 <- survfit(survob~1, conf.type="none") #conf.type="plain" for 95% CI
```

- Compute the support

```{r }
supp1 <- ESPA_supp(time1, status1)
t.grid1  <- seq(supp1[1], supp1[2], by=.1)
```

- Compute survival function over a grid of points

```{r}
saddle.sur1 <- rep(0,length(t.grid1))

for(i in 1:length(t.grid1)){
    saddle.sur1[i] = ESPA_survival(t.grid1[i], time1, status1, inc=.08)
    }

```

- Plot the survival function

```{r}
my.lwd <- 2
title <-  "Survival Function: aml (n=11)"
plot(fit1,
  xlab = "time", 
  ylab = "", 
  main = title,
  lty = c(1,3,3), 
  conf.int = T, 
  yaxs = "i", 
  lwd = my.lwd, 
  xlim = c(supp1[1],supp1[2]), 
  col=c("black","grey","grey"))

points(t.grid1, saddle.sur1, type='l', lwd=my.lwd, col='blue', lty=2)
legend('bottomleft', legend=c('Kaplan-Meier','ESPA'),lwd = my.lwd,
       col=c('black','blue'), lty=c(1,2,4))
```

## Ovarian data
    
- Data

```{r}
d <- ovarian
time <- d$futime
status <- d$fustat
```

- KM

```{r}
survob <- Surv(time=time,event=status,type='right')
fit <- survfit(survob~1,conf.type="none")
```

- Compute the support

```{r}
supp <- ESPA_supp(time,status)
t.grid <- seq(supp[1],supp[2],by=1)
```

- Compute survival function over a grid of points

```{r}
saddle.sur <- rep(0,length(t.grid))
for(i in 1:length(t.grid)){
    saddle.sur[i]  <- ESPA_survival(t.grid[i], time, status, inc=.1)
    }
```

- Plot the survival function

```{r}
title <- 'ovarian (n=26)'
plot(fit,
  lty=c(1,3,3),
  main=title,
  xlab="time",
  conf.int=T, 
  ylab="", 
  yaxs="i", 
  lwd=my.lwd, 
  xlim=c(supp[1],supp[2]), 
  col=c("black","grey","grey"))
points(t.grid, saddle.sur, lwd=my.lwd, type='l', col='blue', lty=2)
legend('bottomleft', legend=c('Kaplan-Meier','ESPA'),lwd = my.lwd,
       col=c('black','blue'), lty=c(1,2,4))
```


##  asaur data (gastricXelox: n=48)

- Data

```{r}
d <- gastricXelox
time <- d$timeWeeks
status <- d$delta
```

- KM

```{r}
survob <- Surv(time=time,event=status,type='right')
fit  <- survfit(survob~1,conf.type="none")
```

- Compute the support

```{r}
supp  <- ESPA_supp(time,status)
t.grid <- seq(supp[1],supp[2],by=1)
```

- Compute survival function over a grid of points

```{r}
saddle.sur <- rep(0,length(t.grid))
for(i in 1:length(t.grid)){
    saddle.sur[i] <- ESPA_survival(t.grid[i], time,status, inc=.1)
    }
```

- Plot the survival function

```{r}
title <- 'gastricXelox (n=48)'
plot(fit,lty=c(1,3,3),
  main=title,
  xlab="time",
  conf.int=T, 
  ylab="",
  yaxs="i",
  lwd=my.lwd,
  xlim=c(supp[1],supp[2]), 
  col=c("black","grey","grey"))
points(t.grid, saddle.sur, lwd=my.lwd, type='l', col='blue', lty=2)
legend('bottomleft', legend=c('Kaplan-Meier','ESPA'),lwd = my.lwd,
       col=c('black','blue'), lty=c(1,2,4))
```

## genfan: n=70
    
- Data

```{r}
d <- genfan
time <- d$hours
status <- d$status 
```

- KM

```{r}
survob <- Surv(time=time,event=status,type='right')
fit <- survfit(survob~1,conf.type="none")
```

- Compute the support

```{r}
supp <- ESPA_supp(time,status)
t.grid <- seq(supp[1], supp[2], by=1)
```

- Compute survival function over a grid of points

```{r}
saddle.sur=rep(0,length(t.grid))
for(i in 1:length(t.grid)){
    saddle.sur[i] <- ESPA_survival(t.grid[i], time,status, inc=.1)
    }
```

- Plot the survival function

```{r}
title <- 'genfan (n=70)'
plot(fit,lty=c(1,3,3),
  main=title,
  xlab="time",
  conf.int=T, 
  ylab="",
  yaxs="i",
  lwd=my.lwd,
  xlim=c(supp[1],supp[2]), 
  col=c("black","grey","grey"))
points(t.grid, saddle.sur, lwd=my.lwd, type='l', col='blue', lty=2)
legend('bottomleft', legend=c('Kaplan-Meier','ESPA'),lwd = my.lwd,
       col=c('black','blue'), lty=c(1,2,4))

```

